---
title: "hw_04:Tidy data and joins"
author: "SL_Ivy"
always_allow_html: true
output:
  html_document: 
    keep_md: yes
  github_document:
    toc: true
    toc_depth: 2
---
## Goal

The goal of this homework is to solidify the data wrangling skills by working some realistic problems in the grey area between data aggregation and data reshaping.
`gapminder` data frame will be used to practice the data reshaping and table joins.



## Setup the working environment for the data analysis
Make sure `gapminder`, `tidyverse`, and `DT` packages are installed.
Load them with the following code:
```{r, warning = FALSE, message = FALSE}
library(gapminder)
library(tidyverse)
library(DT)
```

## Exercise 1: Univariate Data Reshaping

The Exercise 1 and 2 need to use the date frame `gapminder`. It would be better to check the structure of this data frame before we get start.
```{r}
gapminder
```

#### Exercise 1.1: Make a tibble with one row per **year**, and columns for **life expectancy** and **GDP per capita** for *Canada* and *United Kingdom*.
```{r}
gap_wider <- gapminder %>%
  filter(year, country=='Canada' | country =="United Kingdom")%>%
  pivot_wider(id_cols     = year, 
              names_from  = country, 
              values_from = lifeExp) %>%
  as_tibble()
print(gap_wider)
```


#### Exercise 1.2 Producing a scatterplot of life expectancy for *Canada* against that of *United Kingdom*.


```{r}
gap_wider %>% 
  ggplot(aes(x=Canada, y=`United Kingdom`)) +
  geom_point(size=2, color='blue') +
  xlab("Canada Life Expectancy (years)") +
  ylab("Mexico Life Expectancy (years)" ) +
  ggtitle("Life Expectancy of Canada vs. United Kingdom")+
  theme_bw()
```


#### Exercise 1.3 Re-length the data
```{r}
gap_wider %>% 
  pivot_longer(cols = c(-year), names_to  = "country", values_to = "LifeExp") %>%
  DT::datatable()
```


## Exercise 2: Multivariate Data Reshaping

#### Exercise 2.1: Make a tibble with one row per **year**, and columns for **life expectancy** and **GDP per capita** for *Canada* and *United Kingdom*.

```{r}
gap_multiwider <- gapminder %>%
  filter(year, country=='Canada' | country =="United Kingdom")%>%
  pivot_wider(id_cols     = year,
              names_from  = country,
              names_sep = " in ",
              values_from = c(lifeExp, gdpPercap)) %>%
  as_tibble()
print(gap_multiwider)
```


#### Exercise 2.2 Re-length the data
```{r}
gap_multiwider %>% 
  pivot_longer(cols = -year, names_to  = c(".value",  "country"), names_sep = ' in ') %>%
  DT::datatable()
```


## Exercise 3: Table Joins 
* First, we need to load the data for this practice session.
Read in the made-up wedding guestlist and email addresses using the following lines:

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

After load the data, we can check the data by calling their assigned name.
* The guest 
```{r}
guest
```

The email
```{r}
email
```

#### Exercise 3.1: For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

It can be found that the in the `email` tibble, there are mutiple guest sharing one email address, but all the other guests' information stored in the guest tibble are based on the individual guest.
Also, the column name in `email` that stores the guests' name is called *guest* other than *name* in the `guest`.
In order to join two tibbles, the variables in the `email` tibble need to be reshaped and renamed to match the `guest` tibble.
Each name in the same row is separated by ", ", so we can split one row based on this `sep=", ". **Pay attention to the space after the comma**

```{r separate email rows}
email_shaped <- email %>% 
  separate_rows(guest, sep = ", ") %>%
  rename(name = guest)%>%
  as_tibble()
```

To add the email for each guest in the `guest` tibble

```{r}
guest  %>%
  left_join(email_shaped, by = "name")
DT::datatable(guest)
```


#### Exercise 3.2: To figure out the guests that appears in the `email` tibble but no in the guestlist, or say `guest` tibble.

```{r}
email_shaped %>%
  anti_join(guest, by = "name")
```

Those are the three guests that somehow haven't been added to the guestlist.

#### Exercise 3.3: Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r}
guest_new <-guest %>%
  full_join(email_shaped, by = "name")
DT::datatable(guest_new)
```

Here are the new guestlist that contains everyone we have emails for.