---
title: "hw_04:Tidy data and joins"
author: "SL_Ivy"
output: 
  html_document: 
    keep_md: yes
---
## Goal

The goal of this homework is to solidify the data wrangling skills by working some realistic problems in the grey area between data aggregation and data reshaping.
`gapminder` data frame will be used to practice the data reshaping and table joins.



## Setup the working environment for the data analysis
Make sure `gapminder`, `tidyverse`, and `DT` packages are installed.
Load them with the following code:
```{r, warning = FALSE, message = FALSE}
library(gapminder)
library(tidyverse)
library(DT)
```

## Exercise 1: Univariate Data Reshaping

#### Exercise 1.1: Make a tibble with one row per **year**, and columns for **life expectancy** and **GDP per capita** for *Canada* and *United Kingdom*.
```{r}
gap_wider <- gapminder %>%
  filter(year, country=='Canada' | country =="United Kingdom")%>%
  pivot_wider(id_cols     = year, 
              names_from  = country, 
              values_from = lifeExp) %>%
  as_tibble()
print(gap_wider)
```


#### Exercise 1.2 Producing a scatterplot of life expectancy for *Canada* against *that of *United Kingdom*.


```{r}
gap_wider %>% 
  ggplot(aes(x=Canada, y=`United Kingdom`)) +
  geom_point(size=2, color='blue') +
  xlab("Canada Life Expectancy (years)") +
  ylab("Mexico Life Expectancy (years)" ) +
  ggtitle("Life Expectancy of Canada vs. United Kingdom")+
  theme_bw()
```


#### Exercise 1.3 Re-length the data
```{r}
gap_wider %>% 
  pivot_longer(cols = c(-year), names_to  = "country", values_to = "LifeExp") %>%
  DT::datatable()
```


## Exercise 2: Multivariate Data Reshaping

#### Exercise 2.1: Make a tibble with one row per **year**, and columns for **life expectancy** and **GDP per capita** for *Canada* and *United Kingdom*.

```{r}
gap_multiwider <- gapminder %>%
  filter(year, country=='Canada' | country =="United Kingdom")%>%
  pivot_wider(id_cols     = year,
              names_from  = country,
              names_sep = " in ",
              values_from = c(lifeExp, gdpPercap)) %>%
  as_tibble()
print(gap_multiwider)
```


#### Exercise 2.2 Re-length the data
```{r}
gap_multiwider %>% 
  pivot_longer(cols = -year, names_to  = c(".value",  "country"), names_sep = ' in ') %>%
  DT::datatable()
```


## Exercise 3: Table Joins 

Read in the made-up wedding guestlist and email addresses using the following lines:

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

#### Exercise 3.1: For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

```{r}

```

