---
title: "Assignment 02: Explore Gapminder and use dplyr"
author: "Ivy Li"
output: 
  html_document: 
    keep_md: yes
  pdf_document: 
    keep_tex: yes
---
<!---The following chunk allows errors when knitting--->
```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

#### Lading packages for the analysis
```{r, warning=FALSE, message=FALSE}
library(gapminder)
library (tidyverse)
```

## Exercise 1: Basic `dplyr`: Use `dplyr` functions to achieve the following.

#### 1.1 Use `filter()` to *subset* the `gapminder` data to three selected countries in the 1970’s.

* Selected countries: Afghanistan, Albania, and Algeria

```{r, echo=FALSE}
gapminder %>%
  filter(year < 1980 & year >1969, 
         country == 'Afghanistan'| country== 'Albania' | country== 'Algeria') %>%
  knitr::kable(format = 'html', align = 'c')
```

#### 1.2 Use the pipe operator `%>%` to select “country” and “gdpPercap” from the filtered dataset in 1.1

```{r, echo=FALSE}
gapminder %>%
  filter(year < 1980 & year >1969, 
         country == 'Afghanistan'| country== 'Albania' | country== 'Algeria') %>%
  select(country, gdpPercap) %>%
  knitr::kable(format = 'html', align = 'c')
```

#### 1.3 Filter `gapminder` to all entries that have experienced a drop in life expectancy.

```{r}
gapminder %>%
  arrange(year) %>% 
  group_by(country) %>% 
  mutate(diff_LifeExp = lifeExp - lag(lifeExp)) %>% 
  filter(diff_LifeExp < 0) %>% 
  arrange(country) %>%
  knitr::kable(format = 'html', align = 'c')
```

#### 1.4 The max GDP per capita experienced by each country.

```{r}
gapminder %>%
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)) %>% 
  arrange(country) %>%
  knitr::kable(format = 'html', align = 'c') 
```

#### 1.5 A scatterplot of Canada’s life expectancy vs. GDP per capita

```{r}
gapminder %>%
  filter (country == 'Canada') %>%
  ggplot(aes(gdpPercap, lifeExp))+
  geom_point(alpha = 1, colour= 'purple')+
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) +
  theme_classic() +
  ylab("Life Expectancy")
```

## Exercise 2: Explore individual variables with `dplyr`

#### Pick one categorical variable and one quantitative variable to explore.
* Selected variables:
  * categorical variable: continent
  * quantitative variable: lifeExp

### What are the possible values of each variable?
1.1 *Continents* that contained in the `gapminder`:
```{r, echo=FALSE}
gapminder %>%
  select(continent) %>%
  summary()
```

1.2 The distribution of continents:
```{r}
gapminder %>%
  ggplot(aes(continent)) +
  geom_bar(colour = 'pink') +
  theme_classic()+
  xlab("Continent")+
  ylab("Frequency")
```


2.1 The range of the *life expectabct* for `gapminder` data frame
```{r}
gapminder %>%
  select(lifeExp) %>%
  range()
```


2.2 The possible values of the *life expectancy* for the entire data frame:
```{r, echo=FALSE}
gapminder %>%
  select(lifeExp) %>%
  summary()
```
2.3 The mean of the *life expectancy*:
```{r}
mean(gapminder$lifeExp)
```

## Exercise 3: Explore various plot types

* A scatterplot of two quantitative variables in `gapminder`
```{r}
gapminder %>%
  filter(country == 'Canada')%>%
  ggplot(aes(year, lifeExp)) +
  geom_line(colour='pink') +
  theme_classic()+
  xlab("Year")+
  ylab("lifeExp")+
  ggtitle("Changes of life expectancy in Canada from 1952 to 2007")
```

* A scatterplot of two quantitative variables in `mtcars`

The description of the dataset can be found in R-help by running the code below:
```{r}
?mtcars
```
To explore if there is a corelation between car weight and Miles/(US) gallon.
```{r, message=FALSE}
mtcars %>%
  group_by(wt) %>% 
  dplyr::summarise(mpgmean = mean(mpg)) %>%
  plyr::arrange(desc(mpgmean))
mtcars %>%
  group_by(wt) %>%
  ggplot(aes(wt, mpg)) +
  geom_bin2d(colour='pink') +
  theme_classic()+
  xlab("Weight")+
  ylab("Miles/(US) gallon")+
  geom_smooth()+
  ggtitle("Miles/(US) gallon by Weight")
```


