---
title: "Assignment 3"
author: "SL_Ivy"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup
Make sure `gapminder` and `tidyverse` packages are installed.
Load them with the following code:

```{r, message=FALSE, warning=FALSE}
library(gapminder)
library(tidyverse)
```

### Task Option 2
Get the maximum and minimum of GDP per capita for all continents.
```{r}
GDP_summary <- gapminder %>%
  select (continent, gdpPercap) %>%
  group_by(continent) %>%
  summarise(Max_GDP=max(gdpPercap), Min_GDP=min(gdpPercap))
knitr::kable(GDP_summary, format = 'html', align ='c')
```

To obtain a simple figure that only show the Maximum and Minimum values:
```{r}
ggplot(GDP_summary, aes(x=continent)) + 
  geom_jitter(aes(y=Max_GDP, color="Max_GDP")) + 
  geom_jitter(aes(y=Min_GDP, color="Min_GDP"))+
  scale_y_log10(labels = scales::comma_format())+
  ggtitle("Maximum and MInimumGDP Per Capita") 
```

* Or to use the `geom_boxplot` to obtain a more comprehensive figure:
```{r echo=FALSE}
ggplot(gapminder, aes(continent, gdpPercap, group = continent, colour = continent)) +
  geom_boxplot() +
  scale_y_log10(labels = scales::comma_format()) +
  ggtitle("Maximum and MInimumGDP Per Capita")
  
```

### Task option 3

* Look at the spread of GDP per capita within the continents
Generate a table that summarise mean GDP per Cap for each continent:

```{r}
Mean_GDP <- gapminder %>%
  group_by(continent) %>%
  select(continent, gdpPercap)%>%
  summarise_all("mean")
knitr::kable(Mean_GDP, format = 'html', align ='c')
```


* To zoom in to each continent by`subset`:

**The table is too large, it has been hidden, the alternative table can be found in the following _tips_**

For example, Asia: 
```{r, results ='hide'}
Asia_GDP <- subset.data.frame(gapminder, continent == 'Asia')
knitr::kable(Asia_GDP, format = 'html', align ='c')
```


## tips
* Becasue the data table here is too large, the other useful package `kableExtra`can be used to generate a __shorter table__:

```{r, warning=FALSE, message=FALSE}
library(kableExtra)
#and then make the table again with kableExtra#
knitr::kable(Asia_GDP, format = 'html', align ='c')%>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed","response"))%>%
  scroll_box(width = "500px", height = "200px")
```


_All the other continents can be reviewed by the method above by using corresponding name._


* To generate a group of figure to get a overview of the GDP per capita within the continents from 1952 to 2007.
```{r}
ggplot(gapminder, aes(x = year, y = gdpPercap, colour = country)) + geom_line() + scale_y_log10() + facet_wrap(~ continent)+theme(legend.position = "none", axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

### Task option 5
How is life expectancy changing over time on different continents?

1. Find the median and life expectancy for each continent from 1952 to 2007:

```{r}
change_lifeExp <-gapminder %>%
  group_by(continent, year) %>%
  select (continent, year, lifeExp)%>%
  summarize(Median_LifeExp = median(lifeExp), Mean_lifeExp = mean(lifeExp))
knitr::kable(change_lifeExp, format = 'html', align ='c')
```

```{r figure}
# Create a scatter plot showing the change in medianLifeExp over time
ggplot(change_lifeExp, aes(x = year, y = Mean_lifeExp, color = continent)) + geom_line() + scale_x_log10()
```
