---
title: "Assignment 3"
author: "SL_Ivy"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup
Make sure `gapminder` and `tidyverse` packages are installed.
Load them with the following code:
```{r, message=FALSE, warning=FALSE}
library(gapminder)
library(tidyverse)
```


### Task Option 1

Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: how many countries on each continent have a life expectancy less than *27*, for each year.

```{r}
gapminder %>%
  select (lifeExp, year) %>%
  group_by(year) %>%
  arrange (year)

  
```


### Task Option 2
Get the maximum and minimum of GDP per capita for all continents.
```{r}
GDP_summary <- gapminder %>%
  select (continent, gdpPercap) %>%
  group_by(continent) %>%
  summarise(Max_GDP=max(gdpPercap), Min_GDP=min(gdpPercap))
knitr::kable(GDP_summary, format = 'html', align ='c')
```

To obtain a simple figure that only show the Maximum and Minimum values:
```{r}
ggplot(GDP_summary, aes(x=continent)) + 
  geom_jitter(aes(y=Max_GDP, color="Max_GDP")) + 
  geom_jitter(aes(y=Min_GDP, color="Min_GDP"))+
  scale_y_log10(labels = scales::comma_format())+
  ggtitle("Maximum and MInimumGDP Per Capita") 
```

Or can use the `geom_boxplot` to obtain a more comprehensive figure:
```{r echo=FALSE}
ggplot(gapminder, aes(continent, gdpPercap, group = continent, colour = continent)) +
  geom_boxplot() +
  scale_y_log10(labels = scales::comma_format()) +
  ggtitle("Maximum and MInimumGDP Per Capita")
  
```

### Task option 3
* Look at the spread of GDP per capita within the continents
Generate a table that summarise mean GDP per Cap for each continent:






```{r}
Mean_GDP <- gapminder %>%
  group_by(continent) %>%
  select(continent, gdpPercap)%>%
  summarise_all("mean")
knitr::kable(Mean_GDP, format = 'html', align ='c')


```

To zoom in to each continent by`subset`:
For example, Asia: 
```{r}
Asia_GDP <- subset.data.frame(gapminder, continent == 'Asia')
knitr::kable(Asia_GDP, format = 'html', align ='c')
```

