---
title: 'Exercise 2: R Markdown for Gapminder exploration'
author: "Ivy Li"
date: "9/26/2019"
output:
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  github_document:
    toc: yes
---

## About the "gapminder"
`gapminder` is a package created for the purpose of teaching and making code examples. It excerpt from the [Gapminder data](https://www.gapminder.org/data/).
The main object in this package is the `gapminder` data frame.

To explore the data, the `gapminder` package need to be loaded by runing the code below:

```{r}
library(gapminder)
```

## Useful Functions for Exploring Data Frames

* Use `nrow()` and `ncol()` to get the number of rows and number of columns, respectively. This could give you a general picture of the size of the data frame.

```{r,collapse=TRUE}
nrow(gapminder) # show the number of rows
ncol(gapminder) # show the number of columns
```


* `str` is a diagnostic function that compactly display the internal structure of an R object. It returns many useful pieces of information, including the number of observations, varaiables, and the types of data for each column. 
```{r}
str(gapminder)
```

* `summary` is a generic function used to produce result summaries of the results of various model fitting functions. Sometimes, it is the alternative option to `str`. It is essentially applied to each column, and the results for all columns are shown together. 

```{r}
summary(gapminder)
```
